# Practical Machine Learning: Peer Assignment
Machine learning algorithm to predict activity quality from activity monitors using this data set:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

To know how the data was collected read: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Loading data

```{r}
set.seed(1975)
library(caret)
data <- read.csv("pml-training.csv")
```

Removing not used columns. These have data about the person, date and time, and a id.
```{r}
data <- data[,8:160]
inTrain <- createDataPartition(y=data$classe,
	p=0.6, list=F)
tr <- data[inTrain,]
te <- data[-inTrain,]
```


## Create training and testing partitions.
First removing **NA** values and **"#DIV/0!"** values and then creating the training (tr) and testing (te) sets.

```{r}
inTrain <- createDataPartition(y=data$classe,
	p=0.6, list=F)
data[is.na(data)] <- 0
data[data == "#DIV/0!"] <- 0
tr <- data[inTrain,]
te <- data[-inTrain,]
```

## Reducing dimensions in the data set

As 152 variables is a big amount for a complex algorithm like the one used, **random forest**, we look for variables with low variance and remove them.
```{r}
nzv <- nearZeroVar(tr, saveMetrics=T)
tr <- tr[,!nzv$nzv]
```

## Constructing the predictor
The predictor consists in a preproces with **principal component analysis** and a model based in **random forest** bootstraped in 25 times for cross validation.

```{r}
system.time(preProc <- preProcess(tr[,-53],method="pca"))
system.time(trainPC <- predict(preProc, tr[,-53]))
system.time(modelFit <- train(tr$classe ~ ., data=trainPC))
modelFit
```

## Testing the model

The model results with a 0.97 accuracy over the testing set.

```{r}
te <- te[,!nzv$nzv]
testPC <- predict(preProc, te[,-53])
confusionMatrix(te$classe,predict(modelFit,testPC))
```

## Testing with the 20 given cases for that propouse.

```{r}
testing <- read.csv("pml-testing.csv")
testing <- testing[,8:160]
testing <- testing[,!nzv$nzv]
testingPredictions <- predict(preProc, testing[,-53])
answers <- predict(modelFit, testingPredictions)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
